---
title: "INFO7374 - Kiva & World Development Indicators"
author: "Bowei Wang, Dongyue Li, Sarthak Agarwal, Sriram Chandramouli"
date: "7 June 2016"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

# 1. Introduction

A microfinance institution is an organization that is a source of financial services for small businesses lacking access to traditional banking systems. Kiva (www.kiva.org) is a microlending website that works with such microfinance institutions to raise funds for low income entrepreneurs and provide loans for their business. Through this platform, anyone with an internet connection can make an interest free loan.

Kiva has field partners which are mainly microfinance institutions but also schools, NGOs and social enterprises. These field partners work at the local level and disburse loans to the borrowers. Loans are pre-disbursed which means that the loan is given out before being funded on the Kiva website. The field partners then collect stories, photos, and videos from the borrower and post them on Kiva which are published on the website after review. Anyone who can access the Kiva website can browse through the borrower's profile to make a loan for at least 25$. Kiva aggregates all the money and backfills the loan already disbursed by the field partner. Field partners then collect repayments from the borrowers. Some microfinance institutions may charge an interest rate but around 60% of the partners are non-profits. The amount is then repaid to the Kiva lender.

Every year the World Bank releases a collection of World Development Indicators (WDI). The data represents the economic, demographic, social, environmental, educational and cultural indicators. It contains over 800 indicators covering more than 150 countries representing the most current and accurate global development data available which includes national, regional and global estimates.


## Analysis Goals

Kiva's complete business model is based around lending and borrowing. The main focus of Kiva is to be a medium between the loan borrowers and lenders. Hence lenders are very important for Kiva and eventually to loan borrowers. The key objective of this analysis is to help Kiva and loan lenders make better decisions thereby helping in the economic development of the countries and aiding poverty alleviation. In this report we analyze the loan amounts, and their relationship with other variables such as sectors, genders and countries. We also analyze the loan amount fundings and the repayment trends for these loans. 

We further analyze the loans funded across countries with their GDP and provide information about the countries that need immediate attention.

# 2. Data Profile
In the following two sub sections we understand the data and its structure. We further identify the relevant variables within the Kiva dataset that are important for our analysis. 

## 2.1 Dataset description
The Kiva dataset consists of information on lenders and loans. We work primarily with the loans dataset and reference the lenders dataset where necessary. The data is available at www.build.kiva.org website and has three sub directories - `lenders`, `loans` and `loans_lenders`.

```{r, include=FALSE}
library(dplyr)
library(magrittr)
library(RJSONIO)
library(rlist)
library(WDI)
library(ggplot2)
library(parallel)
library(ffbase)
library(reshape2)
library(gridExtra)
library(gapminder)
library(grid)
library(maps)
library(plyr)
library(arules)
```

```{r results='hide', include=FALSE}
data.folder <- "D:/GRAD_SCHOOL/Summer2016/A1/kiva_ds_json/loans/"
loan.file <- list.files(data.folder, pattern="*.json")
```

To analyze the structure, we load the `r length(loan.file)` files from the `loans` sub directory.

## 2.1 Description of rows and observations
Each loan entry in Kiva has the following 30 variables:
```{r, echo=FALSE}
filename = paste(data.folder, loan.file, sep="")
file1.list <- fromJSON(filename[1])
names(file1.list$loans[[1]])
```
For our analysis, we choose the following variables and create a few additional variables:

| Variable            | Type       | Description| 
| ------------------- | ---------- | ----------------------------------------------|
| `location.country`  | Character  | The country of the loan borrower              |
| `CountryF`          | Character  | Country variable converted to factor variable |
| `borrowers.gender`  | Character  | Gender of the borrower                        |
| `borrowers.GenderF` | Character  | Gender variable converted to factor variable  |
| `loan_amount`       | Number     | The amount of loan taken by the borrower      |
| `loan_amount_numeric`| Number    | Country variable converted to factor variable |
| `status`            | Character  | Loan payment status                           |
| `statusF`           | Character  | Status variable converted to factor variable  |
| `sector`            | Character  | Economic sector of the loan entry             |
| `Sector_F`          | Character  | Sector variable converted to factor variable  | 
| `paid_amount`       | Number     | The amount of money paid back by the borrower |
| `posted_date`       | Character  | Date on which the loan is posted on Kiva      |
| `cleanedyear`       | Numeric    | Year extracted from the posted_date           | 
| `YearF`             | Character  | cleanedyear variable converted to factor variable |
| `funded_date`       | Character  | Date on which loan is funded                  |
| `funded_date_cleaned`| Date      | funded_date converted to YYYY-MM-DD format    |
| `paid_date`         | Character  | Date on which loan is paid off                |
| `paid_date_cleaned` | Date       | `paid_date` converted to YYYY-MM-DD format    |
| `day_diff`          | Numeric    | Difference between funded date and paid date  |
| `activity`          | Character  | Reason for which loan is taken                |
| `lender_count`      | Number     | Number of lenders                             |

Units of these variables are:

| Variable      | Unit   |
| --------------|--------|
| `loan_amount` | Dollar |
| `loan_amount_numeric` | Dollar |
| `paid_amount` | Dollar |
| `paid_amount_numeric` | Dollar |


# 3. Dataset preparation

To prepare the data for analysis, we:

  1. load the necessary libraries,
  2. load the Kiva loans dataset into dataframe,
  3. load the WDI dataset into dataframe,
  4. create new variables,
  5. merge Kiva and WDI dataframe

## 3.1 Load libraries
The following libraries are used in this report:

* `dplyr`
* `magrittr`
* `RJSONIO`
* `rlist`
* `WDI`
* `ggplot2`
* `parallel`
* `ffbase`
* `reshape2`
* `gridExtra`
* `gapminder`
* `grid`
* `maps`
* `arules`

## 3.2 Loading the Kiva loans dataset into dataframe 
In the next few steps we load the JSON files under the Kiva `loans` sub directory into R. Since there are more than 1 millions loan entries, we use parallel computing to load the complete data.

We create a function which generates a dataframe from a list.
```{r message=FALSE}
dfrow.from.list = function(aList) { 
  data.frame(rbind(unlist(aList)),
             stringsAsFactors=FALSE)
}
```
We use the above function and create another function to read a JSON file into a dataframe. 
We use `isValidJSON` function to skip invalid JSON files.
```{r}
readJSONFileIntoDataFrame <-
  function (filename) { 
    if(isValidJSON(paste(data.folder, 
          filename,
          sep=""))){
    paste(data.folder, 
          filename,
          sep="") %>%
      fromJSON() %>%
      { .$loans } %>%
      list.select(posted_date, location.country = location$country, sector, loan_amount,
                  funded_date, paid_date, status, lender_count, activity, 
                  borrowers = borrowers[1], paid_amount ) %>%
      lapply(dfrow.from.list) %>% 
      bind_rows()
    }
  }
```

We create a socket cluster which creates a set of copies of R running in parallel. Based on the system configuration we use four cores and use the function `clusterExport` which provides several ways to parallelize computations.
```{r}
cl <- makeCluster(4)
clusterExport(cl,
              c('dfrow.from.list','isValidJSON', 'data.folder', '%>%',
                'list.select','bind_rows','fromJSON',
                'readJSONFileIntoDataFrame'))
```
We use the parallel version of `Lapply` to apply `readJSONFileIntoDataFrame` function on JSON files, then stop the cluster when finished.
```{r warning=FALSE}
create.loan.df.cl = function(cl, loan.file.in) {
  loan.file.in %>%
    { parLapply(cl, ., readJSONFileIntoDataFrame) } %>%
    bind_rows()
}
loan.df = create.loan.df.cl(cl, loan.file[1:40])
stopCluster(cl)
```

## 3.3 Loading the WDI dataset into dataframe
We choose the following Indicator codes from the WDI dataset. We add them to the `indicator.codes` variable.

```{r} 
indicator.codes = c("NY.GDP.MKTP.CD", "NY.GDP.MKTP.KD.ZG","SP.POP.DPND.OL",
                    "SP.POP.DPND.YG","SP.POP.DPND", "SL.TLF.TOTL.IN")
```

We read the WDI data for this code into the `df` dataframe.
```{r}
df <- WDI(indicator = indicator.codes, extra = TRUE)
df$SL.TLF.TOTL.IN
```


## 3.4 Create new variables
Following new variables are created for our analysis.

+ `cleanedyear`
+ `countryF`
+ `loan_amount_numeric`
+ `yearF`
+ `funded_date_cleaned`
+ `paid_date_cleaned`
+ `sector_F`

## `cleanedyear`
We create a numeric variable called `cleanedyear` by extracting year from `posted_date`. This variable is used for merging the Kiva and WDI dataset.
```{r}
cleanedyear <- substr(loan.df[['posted_date']],1,4)
loan.df[["cleanedyear"]] <- as.numeric(cleanedyear)
summary(loan.df[["cleanedyear"]])
```
We observe that the loan data is available from 2006 to 2013 and most of the loans are in the later years.

## `countryF`
We create this variable by converting `location.country` into a factor variable.
```{r}
loan.df$countryF <- factor(loan.df$location.country)
```

## `loan_amount_numeric`
We create this variable by converting the `loan_amount` variable, which is character to numeric.
```{r}
loan.df$loan_amount_numeric <- as.numeric(loan.df$loan_amount)
```

## `yearF`
We create this variable by converting the `cleanedyear` variable into a factor variable.
```{r}
loan.df$yearF <- factor(loan.df$cleanedyear)
```

## `funded_date_cleaned`
We create this variable from the `funded_date` variable by removing unnecessary time information and extracting the date in the format YYYY-MM-DD.
```{r}
loan.df$funded_date_cleaned <- 
  as.Date(substr(loan.df$funded_date, 1, 10), "%Y-%m-%d")
```

## `paid_date_cleaned`
We create this variable from the `paid_date` variable by removing the unnecessary time information and extracting the date in the format YYYY-MM-DD.
```{r}
loan.df$paid_date_cleaned <- 
  as.Date(substr(loan.df$paid_date, 1, 10), "%Y-%m-%d")
```

## `sector_F`
We create this variable by converting `sector` variable into a factor variable.
```{r}
loan.df$sector_F <- factor(loan.df$sector)
```



## 3.5 Merging Kiva and WDI dataframes
We use the `cleanedyear` and `location.country` variables in loan dataframe and join them on the basis of `country` and `year` variables in the WDI dataframe. We use left join to gather all the data from Kiva dataset after merging.
```{r}
finaldf <-
  merge(loan.df, df, by.x = c("location.country", "cleanedyear"),
        by.y = c("country", "year"), all.x = TRUE
  )
```
We remove the `null` records.
```{r}
final1df <- finaldf[!is.na(finaldf$NY.GDP.MKTP.KD.ZG),]
final1df <- finaldf[!is.na(finaldf$SP.POP.DPND),]
final1df <- finaldf[!is.na(finaldf$SP.POP.DPND.OL),]
final1df <- finaldf[!is.na(finaldf$SP.POP.DPND.YG),]
final2df <- finaldf[!is.na(finaldf$SL.TLF.TOTL.IN),]

final2df$SL.TLF.TOTL.IN
```

# 4. Variable summaries and visualizations
In the next few sections, we analyze the significant single and multiple variables for our analysis. 

## 4.1 Total loan amount by country
Our analysis is primarily focuses on the loan amount and its relationship with other variables. 
We visualize the loans and their distribution across various countries by plotting the loan amounts on a world map.  

We load the map data from `world2` and exclude Antarctica from the map.
```{r}
world <- map_data("world2")
world <- subset(world,region!="Antarctica")
```
We calculate the total loan amount for each country.
```{r}
loan_amount_by_country <- condSum(loan.df$loan_amount_numeric,
                                  loan.df$countryF,
                                  na.rm = FALSE)
```
We match the country name with `loan_amount_by_country` variable.
```{r}
world$loanamount <- loan_amount_by_country[
  match(world$region,
        names(loan_amount_by_country),
        nomatch=NA)]
```
We fill the data with the total loan amount by country and move the legend to the bottom of the graph.
```{r}
map <- qplot(long, lat, data = world, 
             group = group, fill = loanamount, 
             geom ="polygon",ylab="",xlab="")
map + theme(legend.position="bottom",
            legend.key.width = unit(3, "line"))
```
We observe that amongst 250 countries for which loans are posted on the Kiva website, maximum loan requirement is in the developing countries of South America, Africa and Asia. Our report focuses on these countries and aims at providing insights that will help in redirecting more funds to these countries. 

## 4.2 Loan requirement by country
We identify the top 10 countries that have the largest loan requirements for the year range 2006-2011. This will help lenders to focus on funding loans for these specific countries thereby catering to their economic development.  

We calculate total loan amount for every country and year.
```{r}
loan_amount_by_country <- 
  condSum(loan.df$loan_amount_numeric,
          list(loan.df$countryF,loan.df$yearF),
          na.rm = FALSE)
```
We create a dataframe of the above matrix.
```{r}
loan_amount_by_country_df <- 
  data.frame(row.names(loan_amount_by_country),
             loan_amount_by_country)
```
We subset the above dataframe for the year range 2006-2011 and rename columns for readability.
```{r}
loan_amount_by_country_df_from2006_to2011 <- 
  loan_amount_by_country_df[,c("row.names.loan_amount_by_country.",
                               "X2006","X2007","X2008","X2009",
                               "X2010","X2011")]
colnames(loan_amount_by_country_df_from2006_to2011) <- 
  c("country", "2006", "2007","2008",
    "2009", "2010","2011")
```
We calculate total loan amount for each country from 2006-2011.
```{r}
loan_amount_by_country_df_from2006_to2011$total <-
  rowSums(loan_amount_by_country_df_from2006_to2011[2:7])
```
We sort this dataframe on the basis of total loan amount and get the records for top 10 countries which have the largest total loan amounts.
```{r}
loan_amount_by_country_df_from2006_to2011_top10 <- head(arrange(
  loan_amount_by_country_df_from2006_to2011, 
               desc(total)), n = 10)
loan_amount_by_country_df_from2006_to2011_top10
```
We observe that Peru, Cambodia and Phillipines are the countries that have the largest loan requirement. Hence lenders can filter the loan postings on Kiva website for these countries.

Also, it is important to note that for Cambodia, Bolivia and Nicaragua loan requirement increases till 2010 and then decreases in 2011. This can be due to economic stagnation or policy changes within the nation.

## 4.3 Loan amount and sector
This analysis provides an understanding of the total loan amount for each sector. There are 15 sectors for which loans are posted on the Kiva website.

We calculate the total loan amount for each sector.
```{r}
loan_amount_by_sector <- condSum(loan.df$loan_amount_numeric,
                                 loan.df$sector_F,na.rm = FALSE)
```
We create the `loan_amount_by_sector_df` dataframe from the `loan_amount_by_sector` variable created above.
```{r}
loan_amount_by_sector_df <-data.frame(names(loan_amount_by_sector),
                                      loan_amount_by_sector)
```
The code chunk below plots the graph for total loan amount for each sector and sorts it on the basis of loan amount.
```{r}
loan_amount_by_sector_df %>%
  ggplot (aes(x = reorder(names(loan_amount_by_sector),
                          loan_amount_by_sector), y = loan_amount_by_sector)) +
  ggtitle("Total Loan amount for each sector") +
  xlab("sectors") +
  ylab("loan amount") +
  geom_bar(stat = "identity") +
  coord_flip()
```
We observe that people are asking for the highest amount of loans in the agriculture, food and retail sectors. Lenders should emphasize on funding loans in these specific sectors. Entertainment and wholesale sectors have the least loan requirement.

To further analyze the total loan amount distribution by sector we use a box plot. 
We modify the wide scale to 10 ~ 90 percentile to eliminate extreme variables. 
```{r warning=FALSE}
loan.df %>%
ggplot(aes(sector_F,loan_amount_numeric)) +
ggtitle("Loan amount distribution by sector") +
xlab("sectors") +
ylab("loan amount") +
geom_boxplot(outlier.shape = NA) +
scale_y_continuous(limits = quantile(loan.df$loan_amount_numeric, c(0.1, 0.9))) +   
coord_flip()
```
In this graph it is interesting to observe that food and retail sector tend to get lower loan amounts inspite we have seen that they get high total loan amount.  

## 4.4 `borrowers.genderF`
The `borrowers.genderF` variable represents the loan borrower's gender.
We create this variable by converting `borrowers.gender` into a factor variable.
```{r}
loan.df$borrowers.genderF <- factor(loan.df$borrowers.gender)
```

The code below prints the summary statistics and a bar graph for visual representation. 
```{r}
summary(loan.df$borrowers.genderF)
loan.df %>% 
  ggplot(aes(x=borrowers.genderF)) +
  geom_bar(aes(fill=borrowers.genderF))
```
We observe that around 70 percent of borrowers on the Kiva website are females.


## 4.5 Loan amount, gender and sector
We analyze the relationship between the total loan amount, gender and sector. Our goal is to visualize the gender based distribution of loan amount in each sector.

Below code plots a mosaic graph which represents the gender on the x-axis, sector on y-axis.
```{r}
mosaicplot(sector_F ~ borrowers.genderF, data = loan.df,
           main = "Gender of borrower for each sector",
           xlab = "gender", ylab = "sector",
           dir = c('h','v'), las = 2,
           col = c(2:9))
```
We observe that 70 percent of the total loans posted are by females. However, some sectors like construction, transportation and manufacturing have more loan postings by males.


## 4.6 Country's GDP and their loan amount requirement
We analyze the relationship between the total loan amount for the countries and the GDP of those countries. As Kiva aims to alleviate poverty in the developing countries, this analysis helps Kiva and the lenders to contribute towards the country's development by making informed decisions on funding loans.

We conduct this analysis for a year range of 2006 to 2011.

We drop all null `posted_date` records from the loans dataframe. We observe that in some records `posted_date` is null but `loan_amount` is not null. We remove the records with null `posted_date` before adding the `loan_amount` by year.
```{r warning=FALSE}
loan.df <- loan.df[!is.na(loan.df$cleanedyear),]
```

We create a dataframe called `GDP_df` with GDP, year and top 10 countries which have the largest loan amount requirement.
```{r}
GDP_df <- data.frame(df$year, df$NY.GDP.MKTP.CD, df$country)
```

We have a dataframe called `loan_amount_by_country_df_from2006_to2011_top10` which has the data related to top 10 countries which have the largest loan amount from 2006 to 2011. We change its data structure by converting all year columns into rows and converting the data to `long` format using `melt()` package.
```{r}
loan_amount_by_country_df_from2006_to2011_top10_long <-
  melt(loan_amount_by_country_df_from2006_to2011_top10[1:7], id="country")
```

We merge `loan_amount_by_country_df_from2006_to2011_top10_long` dataframe which has the top 10 countries with the largest amount of total loan for the year range 2006-2011 and `GDP_df` dataframe created above by `country` and `year`. 
```{r warning=FALSE}
loan_GDP_df <- merge(loan_amount_by_country_df_from2006_to2011_top10_long, 
                     GDP_df, by.x = c("country", "variable"),
                     by.y = c("df.country", "df.year"))
```

We rename the columns of the `loan_GDP_df` for better readability.
```{r}
colnames(loan_GDP_df) <- c("country", "year", "loanamount","GDP")
```

We convert `loanamount` and `GDP` columns into rows using melt function.
```{r}
loan_GDP_df <- melt(loan_GDP_df[1:4], id=c("country","year"))
```

We get the country names and store it in a variable `top10_country_name`.
```{r}
top10_country_name <- 
  loan_amount_by_country_df_from2006_to2011_top10$country %>%
  sapply(as.character)
```
We plot different graphs based on the countries and observe the relationship between GDP and loan amount over the years.
```{r}
title <- top10_country_name[1]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[1])
p1 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p1 = p1+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[2]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[2])
p2 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p2 = p2+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[3]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[3])
p3 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p3 = p3+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[4]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[4])
p4 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p4 = p4+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[5]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[5])
p5 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p5 = p5+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[6]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[6])
p6 =loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p6 = p6+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[7]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[7])
p7 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p7 = p7+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[8]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[8])
p8 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p8 = p8+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[9]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[9])
p9 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p9 = p9+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r echo=FALSE}
title <- top10_country_name[10]
loan_GDP_df_filter <- filter(loan_GDP_df, country %in% top10_country_name[10])
p10 = loan_GDP_df_filter %>%
  ggplot(aes(x=year, y=value, group=variable,
             colour=variable)) +
  geom_line()+
  ggtitle(title) +
  xlab("year") +
  ylab("value")
p10 = p10+theme(legend.position="bottom") + facet_grid(variable ~ ., scales = "free_y")
```

```{r message=FALSE}
grid.arrange(p1, p2, ncol=2)
```
Loan requirement in Peru is increasing with its GDP, however Peru faced economic stagnation between 2008-2009. Loan requirement in Cambodia increases with its GDP till the year 2010 and decreases between 2010-2011. 

```{r}
grid.arrange(p3, p4, ncol=2)
```
We observe that GDP of Phillipines declined between 2008 to 2009, however the requirement of loans constantly increased from 2006 to 2011. Loan requirement in Uganda is increasing with its GDP.  

```{r}
grid.arrange(p5, p6, ncol=2)
```
Loan amount requirement in Bolivia increases with the country's GDP but suddenly decreases between 2009-2011. 
We observe that Tajikistan's requirement for loans increases till 2009 then decreases till 2010 and then again increases in 2011. 

```{r}
grid.arrange(p7, p8, ncol=2)
```
We observe that Kiva loans got popular in Kenya only after strugling between 2006-2009. They skyrocketted between 2009-2011. Kenya's GDP increased between 2006-2011. 
Nicargua's GDP increased till 2008 and then decreased in 2009 and then increased in 2010 and 2011. However, the loans requirement increased till 2010 and then decreased in 2011.
```{r}
grid.arrange(p9, p10, ncol=2)
```

Comparing the countries which have the largest loan requirements we observe that the countries with the lowest GDP are asking for higher loan amounts.


## 4.7 Average loan amount, paid amount and sector
Kiva doesn't guarantee that loans will be paid back but it's important for the lender to have that information before lending money. We analyze the average loan amount against the average paid back amount for every sector.
`Loan amount`, `paid amount` and `sector` are required variables for this analysis. The average value of loan and paid amount for each sector is calculated and represented in a bar graph.

Convert `paid_amount` from character to numeric. Create `sel_finaldf` data frame from `loan.df` by selecting `loan_amount` and `sector`.
```{r warning=FALSE}
loan.df["paid_amount"] <- as.numeric(loan.df$paid_amount)

sel_finaldf <- select(loan.df,paid_amount,loan_amount,sector)
sel_finaldf <- na.omit(sel_finaldf)
```
We calculate the mean value of paid amount and loan amount for each sector.
```{r warning=FALSE}
sel_bysec <- ddply(sel_finaldf, .(sector), summarize, 
                   paid1 = mean(paid_amount),
                   loan1 = mean(as.numeric(loan_amount)))
```
We calculate average loan amount and paid amount by sector.
```{r}
dfm <- melt(sel_bysec, id.vars = c('sector'))
ggplot(dfm, aes(x = factor(sector), y = value, fill = variable) ) +
  geom_bar(stat="identity", position = 'dodge') +
  ggtitle("Average loan amount and paid amount by sector") +
  xlab("sector") +
  ylab("amount") +
  coord_flip()
```
We compare the average loan amount and paid amount for every sector and observe that the education and health sector have largest difference between paid loan amount and total loan amount.


## 4.8 Loan amount, status and sector
To understand the working of loans, status plays an important role. To analyze the status variable we create a factor variable called `statusF` from the `status` character variable. 
```{r}
loan.df$statusF <- factor(loan.df$status)
table(loan.df$statusF)
```
We see that there are 8 loan statuses:

* defaulted - Loans are never paid back after not being paid for 6 months. They are financial loss to the lender of that loan.

* expired - If the loan doesn't get fully funded within 30 days then it "expires".

* funded - Loans that are completely funded are in "funded" state.

* fundraising - These are the loans which are not yet funded. Lenders can only lend money to these loans.

* in_repayment - When the loan is in repayment, it means that the loan has been disbursed to the borrower and they are in process of using the funds.

* paid - These are the loans which are fully paid back by the borrower.

* refunded - Few loans or a portion of it needs to be refunded, those loans maintain "refunded" status.


```{r, include=FALSE}
loan.df %>%
  group_by(statusF,sector_F) %>% tally()
```
We observe that 77 percent of the loans have `paid` status and only 2 percent loans are `default`. This increases the confidence of lenders in Kiva's model.

We further visualize the status of loans based on sector and represent it in a mosaic plot where loan status is on the x axis and sector is on y axis. 
```{r}
op <- par(mar = rep(2, 4))
mosaicplot(sector_F ~ statusF, data = loan.df,
           main = "Status of loans for each sector",
           xlab = "status", ylab = "sector",
           dir = c('h','v'), las = 2,
           col = c(2:9), cex.axis = 0.7)
par(op)
```

We conclude that majority of the loans are paid back, however large number of loans are also in `in_repayment` status.
"Education" and "personal use" sectors have the the largest number of loans in `in_repayment` status.
"Housing" has the largest number of loans in the `expired` status.

## 4.9 The time from funded date to payback date by top 10 loan amount countries
Kiva does not provide any information to lenders about timline of loan repayment, but it's imperative for a lender to have knowledge of the timeline patters for prior loans before lending money. We analyze the time period when a loan is funded and is paid back for the countries which have the largest loan requirements.

We calculate the difference in days by subtracting funded date from paid date and convert the days.
```{r warning=FALSE}
loan.df$day_diff <- 
  as.numeric(loan.df$paid_date_cleaned - loan.df$funded_date_cleaned, "days")
```
We select top 10 countries by loan amount using filter.
```{r}
loan.df_date <-
  filter(loan.df, loan.df$countryF %in% c(top10_country_name))
```
We plot a boxplot and change the scale of the graph to eliminate outliers.
```{r warning=FALSE}
loan.df_date %>%
  ggplot(aes(countryF, day_diff)) +
  ggtitle("Day difference between funded date and payback date") +
  xlab("country") +
  ylab("day difference") +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = quantile(loan.df_date$day_diff, c(0.1, 0.9), na.rm = TRUE)) +
  coord_flip()
```
We observe that most of the loans are paid back within a year. We also conclude that Phillippines has the shortest average payback period whereas Kenya and Cambodia have the highest average payback period among the top 10 borrowers. Therefore, Phillipines is the fastest option to get back the money lended.


# 5. Relationships between variables
In the next few sections we analyze the relationship between several variables using association rules.
 
## 5.1 Relationship between Loan amount and other variables

We compare the loan amount and its relationship with `country`, `sector`, `activity`, `status`, `lendercount` and working age population.

```{r, include=FALSE}
final1df["loan_amount"] <- as.numeric(final1df$loan_amount)
final1df["lender_count"] <- as.numeric(final1df$lender_count)
final1df["SP.POP.DPND.YG"] <- as.numeric(final1df$SP.POP.DPND.YG)
final1df["SP.POP.DPND.OL"] <- as.numeric(final1df$SP.POP.DPND.OL)
```
We divide the loan amount into two categories- "0" and "1" where "0" refers to the loan amount which is less than the average loan amount and "1" refers to the loan amount which is greater than the average loan amount. We store this in the variable `loan_amount_check`.

We group the lender count into 4 quantiles - Q1, Q2, Q3, Q4 and store it in `lender_count_quant`. Q1 denotes lender counts which have values in between 0 and 25% of total lender count, Q2 in between 25% and 50% of total lender counts, Q3 in between 50% and 75% of total lender count and Q4 in between 75% and 100% of total lender count.

Similarly we group the percentage of working age population (young - `SP.POP.DPND.YG` and old - `SP.POP.DPND.OL`) into 4 quantiles.
```{r}
finaldf_loanamount <- final1df %>%
  mutate(loan_amount_check = ifelse(loan_amount >=mean(loan_amount), 1, 0)) %>%
  mutate(lender_count_quant = cut(lender_count, 
                                  quantile (lender_count, c (0, .25, .5, .75, 1)), 
                                  labels=c('Q1','Q2','Q3','Q4'))) %>%
  mutate(working_pop_young = cut(SP.POP.DPND.YG, 
                                 quantile (SP.POP.DPND.YG, c (0, .25, .5, .75, 1)), 
                                 labels=c('Y1','Y2','Y3','Y4'))) %>%
  mutate(working_pop_old = cut(SP.POP.DPND.OL, 
                               quantile (SP.POP.DPND.OL, c (0, .25, .5, .75, 1)), 
                               labels=c('O1','O2','O3','O4'))) %>%
  select(lender_count_quant,status,activity,sector,
         location.country,loan_amount_check,borrowers.gender,working_pop_young,
         working_pop_old)

finaldf_loanamount <- data.frame(finaldf_loanamount)
```            

We convert the variables into factors to apply the association rules.
```{r}
finaldf_loanamount["lender_count_quant"] <- as.factor(finaldf_loanamount$lender_count_quant)
finaldf_loanamount["status"] <- as.factor(finaldf_loanamount$status)
finaldf_loanamount["activity"] <- as.factor(finaldf_loanamount$activity)
finaldf_loanamount["sector"] <- as.factor(finaldf_loanamount$sector)
finaldf_loanamount["location.country"] <- as.factor(finaldf_loanamount$location.country)
finaldf_loanamount["loan_amount_check"] <- as.factor(finaldf_loanamount$loan_amount_check)
finaldf_loanamount["borrowers.gender"]  <- as.factor(finaldf_loanamount$borrowers.gender)
finaldf_loanamount["working_pop_young"]  <- as.factor(finaldf_loanamount$working_pop_young)
finaldf_loanamount["working_pop_old"]  <- as.factor(finaldf_loanamount$working_pop_old)
```

We control the number of results by setting support as 0.01, confidence as 0.5 and maximum of three variables.

We set the rhs value to `loan_amount_check` (0 and 1) and lhs to default. The result is sorted by lift value.
```{r warning=FALSE, results='hide'}
apriori.appearance = list(rhs=c("loan_amount_check=0",
                                "loan_amount_check=1"), default="lhs")
apriori.parameter = list(support=0.01, confidence=0.2, minlen=1, maxlen=3)
rules = apriori(finaldf_loanamount, parameter = 
                  apriori.parameter,appearance =   apriori.appearance)
rules.sorted <- sort(rules, by = "lift")

```

The redundant rules are identified and removed using the below code.
```{r warning=FALSE}
subset.matrix <- is.subset(rules.sorted, rules.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
rules.pruned <- rules.sorted[!redundant]
```
The below rule has the third highest lift value.
```{r warning=FALSE}
inspect(rules.pruned[3])
```
The rule states that Cambodia has high number of lenders which are associated with higher loan amount. 
The support value indicates that in 2 percentage of the total loan amount across all countries, Cambodia has high lenders and high loan amount. The confidence score indicates that 99 percentage of high loan amount in Cambodia have high lender count. The lift value of 2.8 means that this scenario is 2.8 times more likely to occur in data.

```{r warning=FALSE}
inspect(rules.pruned[29])
```
The above rule states that higher loan amount is associated with lesser number of young working population and more number of old working population. 

The support value indicates that 1 percentage of all the countries have high old working population, low young working population and have high loan amount requirement. 
The confidence score indicates that 58 percentage of countries that borrow high loans have high old working population and low young working population.
The lift value means that this scenario is 1.6 times more likely to occur in the data. 


```{r warning=FALSE}
inspect(rules.pruned[31])
```
This rule suggests that the countries having low working population who are young males have high loans. 
The support value indicates that out of all the countries 4 percentage of the countries borrow high loans and have low young male working population.
The confidence score indicates that 57 percentage of the countries that borrow high loans have young male working population.
The lift value of 1.6 means that this scenario is 1.6 times more likely to occur in the data.


```{r warning=FALSE}
inspect(rules.pruned[101])
```
This rule states that the high loan amount is associated with male borrowers. This is interesting because we have analyzed earlier that total percentage of loans borrowed by males are low as compared to females but in this association rule we observe that loan amount borrowed by them is high.
THe support value indicates that out of all loan amounts 11 percentage of the loan amounts are high and are taken by males.
The confidence score indicates that 44 percentage of the males borrow higher loan amount.
The lift value of 1.2 indicates that its 1.2 times more likely to occur in the data.


## 5.2 Relationship between status and other variables

We select `defaulted` and `in_repayment` statuses which represents the loans that are not paid back completely. We analyze the risk that is associated with higher loan amounts and hence filter loans that are above the average value.

We create a new dataframe after filtering `defaulted` and `in_repayment` status and select the required variable for comparison.
```{r}
avgloan_df <- loan.df[loan.df$loan_amount > mean(loan.df$loan_amount),]
finaldf_status <- avgloan_df %>%
  select(status,sector,
            location.country,borrowers.gender) %>%
  filter(status == "defaulted" | status == "in_repayment") 
  
finaldf_status <- data.frame(finaldf_status)
```

We convert all variables to factors to apply association rules.
```{r}
finaldf_status["status"] <- as.factor(finaldf_status$status)
finaldf_status["sector"] <- as.factor(finaldf_status$sector)
finaldf_status["location.country"] <- as.factor(finaldf_status$location.country)
finaldf_status["borrowers.gender"] <- as.factor(finaldf_status$borrowers.gender)
```

We control the number of results by setting confidence and support values to low (0.01) as only one status is predicted for higher values.
We set `defaulted` and `in_repayment` status as rhs and other variables as lhs. The result is sorted by lift value.
```{r warning=FALSE, results='hide'}
apriori.appearance = list(rhs=c("status=defaulted","status=in_repayment"),
                          default="lhs")
apriori.parameter = list(support=0.01,confidence=0.01,minlen=1,maxlen=3)

rules_status = apriori(finaldf_status, parameter = 
                         apriori.parameter,appearance = apriori.appearance)

rules_status.sorted <- sort(rules_status, by = "lift")
```

The redundant rules are identified and removed using the below code.
```{r}
subset.matrix <- is.subset(rules_status.sorted, rules_status.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
rules_status.pruned <- rules_status.sorted[!redundant]
inspect(rules_status.pruned)
```


```{r}
inspect(rules_status.pruned[1])
```
The above rule state that Togo has the high female number of borrowers that are associated with `defaulted` loan status.
The support value indicates that 1 percent of all the loans, borrowers in Togo are females and have `defaulted` loan status.
The confidence score indicates that 97% of female borrowers in Togo have `defaulted` loan status.
The lift value of 7.46 indicate that its 7.4 times more likely to occur in the data.

Similar to Togo, Pakistan also has high female gender borrowers which are associated with `defaulted` loan status.
```{r}
inspect(rules_status.pruned[3])
```
The support value indicates that 1 percent of all the loans, borrowers in Pakistan are females and have `defaulted` loan status.
The confidence score indicates that 63 percent of female borrowers in Pakistan have `defaulted` loan status.
The lift value of 4.90 indicate that its 4.9 times more likely to occur in the data.

```{r}
inspect(rules_status.pruned[10])
```
The above rule states the relationship between loans borrowed by males in agriculture sector and `defaulted` loan status.
The support value indicates that 1 percent of total loans, males take loans in agriculture sector and have `defaulted` loan status.
The confidence score indicates that 15 percent of the loans in agriculture sector are borrowed by males and have `defaulted` loan status.
The lift value of 1.1 indicates that its 1.1 times more likely to occur in the data. 


```{r}
inspect(rules_status.pruned[37])
```
This rule states the realtionship between the male borrowers and `defaulted` loan status.
The support value indicates that 4 percent of the total loans, loans are taken by males and are `defaulted`.
The confidence score indicates that 14 percent of the loans taken by males are in `defaulted` loan status.
The lift value of 1.1 indicates that its 1.1 times more likly to occur in the data.



## 5.3 Relationship between GDP and other variables

We have already described about loan amount distribution across various sectors, genders and countries. We will now observe the relationship between GDP and these variables. We use GDP per annual growth rate for this analysis. This indicator from WDI is already merged with Kiva and the resultant dataframe is `final1df`.

We group GDP values into four quantiles, Q1 is the GDP growth rate in between 0% and 25% of total rates, Q2 is the GDP growth rate in between 25% and 50% of total rates, Q3 is the GDP growth rate in between 50% and 75% of total rates and Q4 is the GDP growth rate in between 75% and 100% of total growth rates.

The loan amount is grouped into four values - "low", "medium", "high" and "very high" based on the value of amount.
```{r}
str(final2df)
final2df["loan_amount"] <- as.numeric(final2df$loan_amount)

#max(final2df$SL.TLF.TOTL.IN)

finaldf_LF <- final2df %>%
  mutate(LF_quant = cut(SL.TLF.TOTL.IN, 
      quantile (SL.TLF.TOTL.IN, c (0, .25, .5, .75, 1)), 
          labels = c('Q1','Q2','Q3','Q4'))) %>%
  mutate(loan_quant = cut(loan_amount, 
      quantile (loan_amount, c (0, .25, .5, .75, 1)), 
          labels=c('low','medium','high','veryhigh'))) %>%
  select(location.country,loan_quant,borrowers.gender,LF_quant)

#finaldf_LF$LF_quant
```

To apply the rules of association, the variables should be converted to factors.

```{r}
finaldf_LF["location.country"] <- as.factor(finaldf_LF$location.country)
finaldf_LF["sector"] <- as.factor(finaldf_LF$sector)
finaldf_LF["borrowers.gender"] <- as.factor(finaldf_LF$borrowers.gender)
finaldf_LF["LF_quant"] <- as.factor(finaldf_LF$LF_quant)
finaldf_LF["loan_quant"] <- as.factor(finaldf_LF$loan_quant)


```

We control the number of results by setting support as 0.01, confidence oas 0.2 and maxlen as 2.

The different quantiles of GDP growth rate are set to rhs value other variables as lhs. The result is sorted by lift value.

```{r warning=FALSE, results='hide'}

apriori.appearance = list(rhs=c("LF_quant=Q1",
      "LF_quant=Q2","LF_quant=Q3", "LF_quant=Q4"),default="lhs")
apriori.parameter = list(support=0.01,confidence=0.2,minlen=1,maxlen=3)
rules_LF = apriori(finaldf_LF, parameter = 
                  apriori.parameter, appearance =  apriori.appearance)
rules_LF.sorted <- sort(rules_LF, by = "lift")

```

The redundant rules are found and removed using the below code

```{r warning=FALSE}
subset.matrix <- is.subset(rules_LF.sorted, rules_LF.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1

rules_LF.pruned <- rules_LF.sorted[!redundant]
inspect(rules_LF.pruned)
```

From our previous statement, it is clear that males borrow higher loans when compared to females. Below rule shows relationship between loan amount and male with respect to GDP.

```{r warning=FALSE}
inspect(rules_GDP.pruned[43])
```

The rule states that in countries where males are associated with higher loans, GDP growth rate is high. This rule has a support value of 0.026% mentioning that 0.026% of males who takes higher loans in any country have higher growth rate with respet to GDP. It has a confidence value of 31% and lift value of 1.28 which indicates that it is 1.28 times more likely to occur in the data.

Another interesting rule with respect to female is described below. We have already stated that females borrow lesser loan amount. Below rule shows this particular relation with respect to GDP.

```{r warning=FALSE}
inspect(rules_GDP.pruned[50])
```

From the above rule, we infer that when females tend to take lesser loans, GDP growth rate of the country seems to be very low(Q1). This has a support value of 0.06 mentioning that 0.06% of females who takes lesser loans in any country have lesser growth rate with respet to GDP. It has a confidence value of 29% and a lift value of 1.16.


# 6. Conclusion




